How to select predictive models for decision-making or causal inference?

### 2023-12-08 how_to_select_causal_models.tex
move prior work

\twocolumn
\section{Background}\label{sec:intro}

@@

\paragraph{Objectives and structure of the paper}

We study \textit{model selection procedures}
in practical settings: \textit{finite samples} settings and without
\textit{well-specification} assumption. Asymptotic causal-inference
theory calls for complex risks, but a practical question is
whether model-selection procedures, that rely on data split, can estimate
these risks reliably enough. Indeed, they
come with more quantities to estimate, which may
bring additional variance, leading to worse model selection.

We first illustrate the problem of causal model selection and briefly review
prior art. Then we anchor causal model selection in the \emph{potential
    outcome} framework and details the causal risks and model-selection procedure.
We then rewrite the so-called $R\text{-risk}$, as a reweighted version of mean
squared difference between the true and estimated individualized treatment
effect. Finally, we conduct a thorough empirical study comparing the different
metrics on a wide variety of data-generative processes based on simulations and
real data.

@@

\subsection{Prior work: model selection for outcome modeling (g-computation)}%

%\section{Formal setting of causal inference and model selection}\label{sec:framework}
\section{Methods}\label{sec:framework}
\subsection{Neyman-Rubin Potential Outcomes framework}%
\label{sec:neyman_rubin}%

@@

Drawing inspiration from the TMLE literature that uses combination of flexible
machine learning methods \cite{schuler_targeted_2017}, we use a meta-learner as models
for the nuisances $\check e$ (respectively $\check m$): a
stacked estimator of ridge and boosting classifiers (respectively
regressions). We select hyper-parameters with randomized search on a validation set
$\mathcal{V}$ and keep them fix for model selection
(\ref{apd:experiments:nuisances_hp} lists hyperparameters).

@@

%\idea{Overlap is important but not easily measurable}
Good overlap between treated and control population is crucial for causal
inference (positivity assumption \ref{assumption:overlap}). We introduce the
Normalized Total Variation (NTV), a divergence based on the propensity score
summarizing the overlap between both populations (Appendix
\ref{apd:motivation_ntv}).

@@

@@

The nuisances are estimated with a stacked regressor inspired by the Super
Learner framework, \cite{laan_super_2007}). The hyper-parameters are
optimized with a random search with following search grid detailed in Table
\ref{apd:experiments:nuisances_hp_grid}. All implementations come from
\href{https://scikit-learn.org/stable/}{scikit-learn}
\cite{pedregosa_scikitlearn_2011}.  As extreme inverse propensity weights induce
high variance, we use clipping \cite{swaminathan_counterfactual_2015,
    ionides_truncated_2008} to bound $min(\check e, 1-\check e)$ away from 0 with a
fixed $\eta=10^{-10}$, ensuring strict overlap for numerical stability.

>>>>

\twocolumn
\section{Introduction}\label{sec:intro}

@@

\section{Objectives and structure of the paper}

We study \textit{model selection procedures} in practical settings:
\textit{finite samples} without the \textit{well-specification}
assumptions which are often made in the literature (Appendix
\ref{apd:prior_work}). Asymptotic causal-inference theory calls for complex
risks, but a practical question is whether model-selection procedures, that rely
on data split, can estimate these risks reliably enough. Indeed, they come with
more quantities to estimate, which may bring additional variance, leading to
worse model selection.

We first illustrate the problem of causal model selection and briefly review
prior art. Then we anchor causal model selection in the \emph{potential outcome}
framework and details the causal risks and model-selection procedure. We then
rewrite the so-called $R\text{-risk}$, as a reweighted version of mean squared
difference between the true and estimated individualized treatment effect.
Finally, we conduct a thorough empirical study comparing the different metrics
on a wide variety of data-generative processes based on simulations and real
data, going beyond prior work
\citep{schuler_comparison_2018,alaa_validating_2019} (reviewed in Appendix
\ref{apd:prior_work}).

@@

% A natural way to select a predictive model for causal inference would be
% an error measure between a causal quantity such as the CATE and models' estimate. But such error is
% not a ``feasible'' risk: it cannot be computed solely from observed data
% and requires oracle knowledge.

\section{Methods}\label{sec:framework}
\subsection{Neyman-Rubin Potential Outcomes framework}%
\label{sec:neyman_rubin}%

@@

Drawing inspiration from the TMLE literature that uses combination of flexible
machine learning methods \cite{schuler_targeted_2017}, we model the nuisances
$\check e$ (respectivley $\check m$) with a meta-learner: a stacked estimator of ridge
and boosting classifiers (respectively regressions) (hyperparameter selection in Appendix \ref{apd:experiments:nuisances_hp}).

@@

%\idea{Overlap is important but not easily measurable}
Good overlap between treated and control population is crucial for causal
inference (Assumption \ref{assumption:overlap}). We introduce the
Normalized Total Variation (NTV), a divergence based on the propensity score
summarizing the overlap between both populations (Appendix
\ref{apd:motivation_ntv}).

@@

\section{Prior work : model selection for outcome modeling (g-computation)}\label{apd:prior_work}

@@

The nuisances are estimated with a stacked regressor inspired by the Super
Learner framework, \cite{laan_super_2007}). We select hyper-parameters
with randomized search on a validation set $\mathcal{V}$ and keep them fix for
model selection  The searc grid is detailed in Table
\ref{apd:experiments:nuisances_hp_grid}. All implementations come from
\href{https://scikit-learn.org/stable/}{scikit-learn}
\cite{pedregosa_scikitlearn_2011}.  As extreme inverse propensity weights induce
high variance, we use clipping \cite{swaminathan_counterfactual_2015,
    ionides_truncated_2008} to bound $min(\check e, 1-\check e)$ away from 0 with a
fixed $\eta=10^{-10}$, ensuring strict overlap for numerical stability.
